---
title: "Coursera_Reproducible_Reasearch_Project"
author: "Lukasz Sadalski"
date: "17 03 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project

This is assignment from week 2 of Reproducible Research Coursera course.

### Loading data and preprocessing

```{r, message = FALSE}

library(dplyr)
library(ggplot2)
library(timeDate)
library(lattice)
```

```{r}

setwd("/Users/lsadalski/Desktop/R/Coursera_Assignment_Report_knitr/RepData_PeerAssessment1")
file <- unzip ("activity.zip")
raw_data <- read.csv("activity.csv")
# str(rdata)
# head(rdata)
# tail(rdata)

raw_data$date <- as.Date(raw_data$date)

```

### What is mean total number of steps taken per day?


```{r}
group_data <- group_by (raw_data, date)
group_sum <- summarise(group_data, sum(steps))
steps <- group_sum$`sum(steps)`
hist (steps, breaks=50, col="green", main = "Total number of steps taken each day", xlab = "Number of steps")
rug(steps)

meansteps <- mean (steps, na.rm = T)
meansteps
mediansteps <- median (steps, na.rm = T)
mediansteps

```

The mean steps per day is `r as.character(meansteps)`
The median steps per day is `r as.character(mediansteps)` 

### What is the average daily activity pattern?

``` {r}
agg_data <- aggregate (raw_data$steps, by = list(raw_data$interval), FUN="mean", na.rm = TRUE)
colnames (agg_data) <- c("interval","average_steps")
ggplot (data=agg_data, aes(x=interval, y=average_steps)) +
  geom_line () +
  xlab ("Interval - 5 minutes") +
  ylab ("Average number of steps") +
  ggtitle ("Average daily activity pattern")
```

Which interval on average across all the days contains maximum number of steps?

```{r}
which_interval <- which.max(agg_data$average_steps)
max_interval <- agg_data[which_interval,1]
max_interval
max_steps <- agg_data[which_interval,2]
max_steps
```

Maximum number of steps is contained in `r max_interval` interval, there was `r max_steps` steps

### Imputing missing values

1. Calculating total number of missing values in the dataset

```{r}
sum(is.na(raw_data$steps))
```

Total number of missing values for steps is `r sum(is.na(raw_data$steps))`

2. Strategy for missing values 

I will replace NA with mean of this 5-minute interval 

3. Creating new data set and replacing NAs

```{r}

imputed_data <- raw_data

for (i in 1:length(imputed_data[, 1])){
    if (is.na(imputed_data[i,1]))
        {  which_interval <- imputed_data[i,3]
           indeks <- which(agg_data$interval == which_interval)
           imputed_data[i,1]<-agg_data[indeks, 2]}
}
```

4. What is mean total number of steps taken per day for imputed data? 

```{r}

group_data <- group_by (imputed_data, date)
group_sum <- summarise(group_data, sum(steps))
steps <- group_sum$`sum(steps)`
hist (steps, breaks=50, col="green", main = "Total number of steps taken each day", xlab = "Number of steps")
rug(steps)

i_meansteps <- mean (steps, na.rm = T)
i_meansteps
i_mediansteps <- median (steps, na.rm = T)
i_mediansteps

```

The mean steps per day is `r as.character(i_meansteps)`
The median steps per day is `r as.character(i_mediansteps)` 

The difference between mean and median for raw and imputed data set (NAs replaced by mean) is for median only.
It was `r as.character(mediansteps)` and now it is `r as.character(i_mediansteps)`

### Are there differences in activity patterns between weekdays and weekends?

```{r}
imputed_data <- mutate (imputed_data,isWeekday(date))
imputed_data$`isWeekday(date)` <- gsub("TRUE","weekday",imputed_data$`isWeekday(date)`)
imputed_data$`isWeekday(date)` <- gsub("FALSE","weekend",imputed_data$`isWeekday(date)`)
imputed_data$`isWeekday(date)` <- as.factor(imputed_data$`isWeekday(date)`)
colnames (imputed_data) <- c("steps","date","interval","weekend_weekday")

agg2_data <- aggregate (imputed_data$steps, by = list(imputed_data$interval, imputed_data$weekend_weekday), FUN="mean", na.rm = TRUE)
colnames (agg2_data) <- c("interval","weekend_weekday","steps")
xyplot(steps~interval|weekend_weekday, agg2_data,type="l",layout=c(1,2), xlab="Interval - 5 minutes", ylab="Average number of steps", main = "Differences in average number of steps on weekend vs weekdays")


```
